Interpretable Neural Networks using EAGGA

### 2025-03-10 paper/main.tex
more bullets on interpretability

% feels better in front of AutoML, as it feels more qualitative, less quantitative
\subsection{Interpretability}
- Why Interpretability is desirable?
- Practical implementations
  (1) post-hoc methods
    - give examples
    - why are they not perfectly suited
  (2) model-based

>>>>

% feels better in front of AutoML, as it feels more qualitative, less quantitative
\subsection{Interpretability}
As there is no clear definition for interpretability, we will consider it as ``the ability to provide \textit{explanations} in \textit{understandable terms} to a human'',
where \textit{explanations} are logical decision rules and \textit{understandable terms} relate to commonly used terms in the domain of the problem,
as suggested by \citet[chap. 1]{survey_NN_interpretability}.
Further, we use the term ``explanability'' in an exchangeable manner with ``interpretability'', as is commonly done.

(Why Interpretability is desirable?)
- in many ways desirable, as per \citet[pp. 3-4]{Zach2019InterpretabilityOD}, e.g.
  * \textbf{gaining trust}: most notable use case, e.g. medical diagnosis: predictions need to be validated by doctors \citep[p. 1]{review_NN_interpretability}
  * \textbf{model debugging}
    - model is optimised w.r.t. loss and judged on accuracy
    - might have unexpected performance drops in certain situations, makes model unreliable \citep[1B]{survey_NN_interpretability}
    - having an interpretable model + model induction process can help find ways to improve accuracy + reliability
  * \textbf{scientific understanding}: when only models can make sense of increasingly complex data anymore, interpretability enables extraction of learnt knowledge encoded
    in the model to make it accessible + reliable to humans (also mentioned here \citep[p. 1]{review_NN_interpretability})
  * \textbf{subconcious bias}, e.g. loan approval: must ensure that decision is not discriminatory
  * \textbf{regulatory}
    - such as EU's "Right to Explanation" warranted by the GDPR \citep[p. 1]{review_NN_interpretability} or
    - drug approval processes in situations where a machine learning model discovered a new drug, process needs to be transparent for the regulator to approve \citep[1B]{survey_NN_interpretability}

Commonly, methods for model interpretation are divided into intrinsic methods, where the search space only comprises models with a structure simple enough to be
considered ``explainable'' (such as tree-based or simple linear models), and post-hoc methods, where interpretation methods are applied after model training.
Amongst post-hoc techniques, we can further divide the space into model-specific (such as analysing GLM coefficients) and model-agnostic
(e.g. partial depence plots, ALE) methods. \citet[chap. 3.2]{molnar2022}

(Taxonomy)
\citet[chap. 2]{survey_NN_interpretability} extend this distinction to a 3-dimensional taxonomy, allowing for better categorisation of neural networks (NN), a model class
that in its regular feedforwad form is inherently non-interpretable.
% alternative
%As regular feedforward neural networks (NNs) are inherently non-interpretable models, the taxonomy introduced by \citet[chap. 2]{survey_NN_interpretability} proves
%particulary useful to categorise NN explanability methods along 3 dimensions:
- \textbf{Passive vs Active Approaches}
  * passive: post-hoc
  * active: changing architecture or training process to make model interpretable
- \textbf{Type of Explanations}
  * examples: providing examples of what leads to desired output
  * attribution: attributing effect on output for a specific feature
  * hidden semantics: examine what type of inputs specific neurons / layers pick up on
  * rules: logical rules, e.g. if-then, trees
- \textbf{Local vs Global Interpretability}
  * local: explanation based on individual samples
  * semi-local: explanation based on set of samples, e.g. grouped together by some criterion
  * global: explaining network as a whole
